Methodology of Creating SVM Kernels from Scratch Using Python and NumPy

### 2023-12-15 Paper/paper.tex
Methodology Section Complete

\documentclass[12pt]{article}
\usepackage{hyperref}
\usepackage{authblk}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{enumitem}
\usepackage{amsfonts}
\usepackage{array}
\usepackage{xcolor}
\usepackage{tikz}
\usepackage{pifont}
\usepackage{fontawesome5}
\usepackage{tikz-3dplot}
\usepackage{forest}
\usepackage{cancel}
\usepackage{ulem}
\usepackage{booktabs}
\usepackage[margin=1in]{geometry}

@@

Support Vector Machines (SVMs) are machine learning algorithms that are used to complete classification tasks. When we have a large dimensional dataset that we would like to create decision boundaries for, we can create a hyperplane that maximizes the margin between the observations' labels. The data points that are closest to the decision boundary influence the position and direction of the hyperplane, and these close data points are called \textbf{Support Vectors}. The goal is to optimize the distance between the support vectors and the hyperplane, and so we can draw a line between the points.

@@

We can combine this all to formulate our optimization problem: minimize $||w||$ constrained by $y_i (w \cdot x_i + b) \geq 1$ for $i = 1,...,N$
We can set up a Lagrange dual optimization problem, such that:
$$L = \frac{1}{2}||w||^2 - \sum_{i=1}^{n} \alpha_i \cdot [y_i(w \cdot x + b) - 1]$$
Minimizing for $w$ and $b$, we get:
$$w = \sum_{i=1}^{n} \alpha_i y_i x_i$$
$$\sum_{i=1}^{n} \alpha_i y_i = 0$$

Our problem 
\subsection{Kernel Functions}
We implemented four different kernel functions which are listed below.
\begin{itemize}
    \item Linear Kernel
    \begin{equation}
    K(X, Y) = X^T Y
    \end{equation}

@@

    \item Radial Basis Function (RBF) Kernel
    \begin{equation}
    K(X, Y) = \exp(-\gamma ||X - Y||^2), \gamma > 0
    \end{equation}
\end{itemize}
\subsection{Binary SVM}
Discuss the implementation of the Binary SVM class, including the implementation of the different kernels.

\subsection{Multiclass SVM}
Discuss the implementation of the Multiclass SVM class, which uses the Binary SVM class.

>>>>

\documentclass[12pt]{article}
\documentclass[12pt]{article}
\usepackage{hyperref}
\usepackage{authblk}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{enumitem}
\usepackage{amsfonts}
\usepackage{array}
\usepackage{xcolor}
\usepackage{tikz}
\usepackage{pifont}
\usepackage{fontawesome5}
\usepackage{tikz-3dplot}
\usepackage{forest}
\usepackage{cancel}
\usepackage{ulem}
\usepackage{booktabs}
\usepackage[margin=1in]{geometry}

@@

Support Vector Machines (\textbf{SVMs}) are machine learning algorithms that are used to complete classification tasks. When we have a large dimensional data set that we would like to create decision boundaries, we can create a hyperplane that maximizes the margin between the observations' labels. The data points that are closest to the decision boundary influence the position and direction of the hyperplane, and these close data points are called \textbf{Support Vectors}. The goal is to optimize the distance between the support vectors and the hyperplane (also known as the \textbf{margin}), and so we can draw a line between the points.

@@

We can combine this all to formulate our optimization problem: minimize $||w||$ constrained by $y_i (w \cdot x_i + b) \geq 1$ for $i = 1,...,N$.
We can set up a Lagrange dual optimization problem, such that:
$$L = \frac{1}{2}||w||^2 - \sum_{i=1}^{n} \alpha_i \cdot [y_i(w \cdot x + b) - 1]$$
Minimizing for $w$ and $b$, we get:
$$w = \sum_{i=1}^{n} \alpha_i y_i x_i$$
$$\sum_{i=1}^{n} \alpha_i y_i = 0$$

Our problem can now be in terms of a constrained minimization problem by plugging in the constraints we got. This can be simplified into a quadratic form. The final form of our problem will be

\begin{equation}
    L = 1^T \alpha - \frac{1}{2}\alpha^TX^TX\alpha
\end{equation}

\subsection{Kernel Functions}
We implemented four different kernel functions which are listed below.
\begin{itemize}
    \item Linear Kernel
    \begin{equation}
    K(X, Y) = X^T Y
    \end{equation}

@@

    \item Radial Basis Function (RBF) Kernel
    \begin{equation}
    K(X, Y) = \exp(-\gamma ||X - Y||^2), \gamma > 0
    \end{equation}
\end{itemize}

The kernel tricks used to form a linear decision boundary on a non-linear data set will be implemented into the optimization problem (1). For the linear kernel, no difference is made because it is the same thing that is done in the equation, just the product of the inputs. However, for the other kernels, the trick is implemented in the transformation of the input:
$$(X^TX)_{ij} = \sum_{d} y_i x_{id} \cdot y_j x_{jd} = \sum_{d} y_i y_k \cdot x_{id} x_{jd}$$
$$(k^Tk)_{ij} = \sum_{d'} y_i k_{id'} \cdot y_j k_{jd'} = \sum_{d'} y_i y_k \cdot k_{id'} k_{jd'}$$

The $k$ is found using the different kernel equations [(2)-(5)], which allows for versatility in the shape of the boundary line. It is the optimal decision boundary that contains

\subsection{Binary SVM}
Discuss the implementation of the Binary SVM class, including the implementation of the different kernels.

A binary classification on the data is the foundation of the calculations. We want to form a hyperplane that separates two labels effectively, with a margin for error when making predictions on new observations. The steps to complete to find the terms for our hyperplane can be seen below:

\begin{itemize}
    \item For each training example $(\mathbf{x}_i, y_i)$:
    \begin{itemize}
        \item Calculate error $E_i$ for $\mathbf{x}_i$ using the current model:
        \begin{equation*}
            E_i = f(\mathbf{x}_i) - y_i
        \end{equation*}
        \begin{equation*}
            f(x_i) = \sum_{j=1}^n \alpha_i y_i K (x_j,x_i) + b
        \end{equation*}
        \item If $(y_i \cdot E_i < -tol \text{ and } \alpha_i < C)$ or $(y_i \cdot E_i > tol \text{ and } \alpha_i > 0)$:
        \begin{itemize}
            \item Select a second example $(\mathbf{x}_j, y_j)$ to update alpha (using heuristics to choose $j \neq i$)
            \item Calculate error $E_j$ for $\mathbf{x}_j$
            \item Save old alpha values: $\alpha_i^{old} = \alpha_i$, $\alpha_j^{old} = \alpha_j$
            \item Compute bounds $L$ and $H$ for $\alpha_j$ (satisfying $0 \leq \alpha_j \leq C$)
            \item If $L$ equals $H$, continue to the next training example
            \item Compute $\eta$ (second derivative of the SVM objective function)
            \item Update $\alpha_j$:
            \begin{align*}
                \alpha_j &= \alpha_j^{old} - \frac{y_j \cdot (E_i - E_j)}{\eta} \\
                &\text{(Clip $\alpha_j$ if necessary: $\alpha_j = \text{clip}(\alpha_j, L, H)$)} \\
                &\text{(Continue if $|\alpha_j - \alpha_j^{old}| < \text{epsilon}$)}
            \end{align*}
            \item Update $\alpha_i$:
            \begin{equation*}
                \alpha_i = \alpha_i^{old} + y_i \cdot y_j \cdot (\alpha_j^{old} - \alpha_j)
            \end{equation*}
            \item Update bias term $b$ (compute $b_1$ and $b_2$, adjust $b$ based on $\alpha_i$ and $\alpha_j$, update $E$ for all examples)
        \end{itemize}
    \end{itemize}
\end{itemize}

Using this algorithm, you are able to find the $w$'s and $b$ that define the hyperplane between the two classes. The $K()$ function represents the kernel that is being used, and so this is where the different transformations will have an effect.

\subsection{Multiclass SVM}
The multi-classification on a data set using SVM is only a slight variation from the binary classification. In binary, we trained the model on a \textbf{1 vs 1} basis. We had two classes, -1 and 1, and wanted to create a decision boundary between them. However, multiple classes means that we have more than two classes, and you can't just draw one line to separate the labels.

Therefore, the meta-algorithm for multi-class is to approach with a \textbf{one vs others} mindset. For however many classes you have, you want to create a decision boundary between one class and all the others, and repeating this until you have done so for each label. It follows the same steps as binary classification, but does it $s$ times for the number of labels in the data set to differentiate.
